% Created 2021-05-03 Mon 04:01
% Intended LaTeX compiler: pdflatex
\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\documentclass{article}
\usepackage{here}
\usepackage{xcolor}
\usepackage[margin=3.0cm]{geometry}
\usepackage{amsmath}
\usepackage{parskip}
\renewcommand\arraystretch{1.4}
\usepackage[margin=1in]{geometry}
\usepackage{minted}
\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newminted{c}{frame=single}
\author{Fatih Kaan Salgır - 171044009}
\date{}
\title{CSE344 - System Programming - Midterm}
\hypersetup{
 pdfauthor={Fatih Kaan Salgır - 171044009},
 pdftitle={CSE344 - System Programming - Midterm},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.4.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Design Explanation}
\label{sec:org77741c8}

\subsection{Children processes}
\label{sec:org14c3797}
The parent process will fork into 3 types of children:

\subsubsection{Nurses}
\label{sec:orga1d9756}
Nurses are the producers. Just like any other producer \& consumer problem they will wait for empty space when the buffer is full. However posting the semaphore to notify consumer is a little bit different. The value of semaphore must be the minimum of 2 vaccines. In order to satisfy this, semaphore should be posted only if the type of the vaccine added to buffer is less than or equal to the other type of vaccine:

\begin{ccode}[fontsize=\small,frame=single,framesep=3mm]{text}
if (!mem->eof && novacc_this <= novacc_other)
  sem_post(&mem->vaccines_available);
\end{ccode}

Nurses will continue until, one of them reach the end of the file, and return value of \texttt{read()} is goint to be 0. In this case nurse will be terminated. But before terminaton, it will perform \texttt{sem\_post(\&mem->empty)} to make sure that the other nurses won't stuck in \texttt{sem\_wait(\&mem->empty)}, and they will also terminate.

\subsubsection{Vaccinators}
\label{sec:org782ef75}
Vaccinators are the consumers. They will wait for the semaphore: \texttt{vaccines\_available}, and post \texttt{empty}. In the critical section:
\begin{enumerate}
\item remove vaccines from the buffer
\item selct the next citizen to invite
\item set the information to print
\end{enumerate}

After critical section, the vaccinator is going to inform the citizen by writing to its pipe. And print the log.

Note: Since printing is out of the critical section, the order of the lines are not guaranteed. I have added both on the Running \& Test Cases section.
\subsubsection{Citizens}
\label{sec:org207eaba}
Citizens should wait for get vaccinated. Thus, a citizen will read its pipe \texttt{t} times (number of shots) in total. The write end of the pipes are located in the shared memory, and they are accesed by vaccinators.

\newpage

\subsection{Number of Semaphores}
\label{sec:org22e9237}
3 unnamed POSIX semaphores are used, and they are located in shared memory.
\begin{itemize}
\item \texttt{sem}: to protect access to shared memory
\item \texttt{empty}: nurses (producers) will wait on this semaphore, when the buffer is full
\item \texttt{vaccines\_available}: vaccinators (consumers) will wait on this semaphore, until there are enough amount of vaccines in the buffer
\end{itemize}

\subsection{Shared Memory}
\label{sec:orgd862a9c}
Multiple unnamed shared memories is used for interprocess communication.  \texttt{mmap} system call is used with \texttt{MAP\_SHARED | MAP\_ANONYMOUS} flags, therefore the mapping is not backed by any file.

\subsection{Handling SIGINT}
\label{sec:org42d912d}
In case of receiving \texttt{SIGINT}, handler function is going to call \texttt{exit(EXIT\_SUCCESS)}. An exit handler is registered, therefore it will terminated gracefully.

\section{Running \& Test Cases}
\label{sec:org6d58fe7}

Program tested with different cases and also with the valgrind to check memory leaks. According the output of the valgrind, "All heap blocks were freed -- no leaks are possible".

\begin{minted}[fontsize=\small,frame=single,framesep=3mm]{text}
valgrind --leak-check=full \
         --show-leak-kinds=all \
         --track-origins=yes \
         --verbose \
         --log-file=valgrind-out.txt \
         ...
\end{minted}

\newpage

When log functions are inside of the critical section:

\begin{minted}[fontsize=\small,frame=single,framesep=3mm]{text}
./program  -n 3 -v 3 -c 3 -b 11 -t 3 -i input
\end{minted}


\begin{minted}[breaklines,fontsize=\footnotesize,frame=single,framesep=3mm]{text}
Welcome to GTU344 clinic. Number of citizens to vaccinate c=3 with t=3 doeses.
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 1 vaccine1 and 0 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 2: the clinic has 1 vaccine1 and 1 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 2 vaccine1 and 1 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 2: the clinic has 2 vaccine1 and 2 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 3 vaccine1 and 2 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 2: the clinic has 3 vaccine1 and 3 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 4 vaccine1 and 3 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 2: the clinic has 4 vaccine1 and 4 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 5 vaccine1 and 4 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 2: the clinic has 5 vaccine1 and 5 vaccine2.
Nurse 2 (pid=861724) has brought vaccine 1: the clinic has 6 vaccine1 and 5 vaccine2.
Vaccinator 1 (pid=861729) is inviting citizen pid=861726 to the clinic.
Citizen 1 (pid=861726) is vaccinated for the 1st time: the clinic has 5 vaccine1 and 4 vaccine2. 
Vaccinator 2 (pid=861730) is inviting citizen pid=861727 to the clinic.
Citizen 2 (pid=861727) is vaccinated for the 1st time: the clinic has 4 vaccine1 and 3 vaccine2. 
Nurse 3 (pid=861725) has brought vaccine 2: the clinic has 4 vaccine1 and 4 vaccine2.
Vaccinator 3 (pid=861731) is inviting citizen pid=861728 to the clinic.
Citizen 3 (pid=861728) is vaccinated for the 1st time: the clinic has 3 vaccine1 and 3 vaccine2. 
Vaccinator 3 (pid=861731) is inviting citizen pid=861726 to the clinic.
Citizen 1 (pid=861726) is vaccinated for the 2nd time: the clinic has 2 vaccine1 and 2 vaccine2. 
Vaccinator 2 (pid=861730) is inviting citizen pid=861727 to the clinic.
Citizen 2 (pid=861727) is vaccinated for the 2nd time: the clinic has 1 vaccine1 and 1 vaccine2. 
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 2 vaccine1 and 1 vaccine2.
Nurse 1 (pid=861723) has brought vaccine 2: the clinic has 2 vaccine1 and 2 vaccine2.
Nurse 3 (pid=861725) has brought vaccine 1: the clinic has 3 vaccine1 and 2 vaccine2.
Vaccinator 1 (pid=861729) is inviting citizen pid=861728 to the clinic.
Citizen 3 (pid=861728) is vaccinated for the 2nd time: the clinic has 2 vaccine1 and 1 vaccine2. 
Nurse 2 (pid=861724) has brought vaccine 2: the clinic has 2 vaccine1 and 2 vaccine2.
Vaccinator 3 (pid=861731) is inviting citizen pid=861726 to the clinic.
Citizen 1 (pid=861726) is vaccinated for the 3rd time: the clinic has 1 vaccine1 and 1 vaccine2. The citizen is leaving. Remaining citizens to vaccinate: 2
Nurse 1 (pid=861723) has brought vaccine 1: the clinic has 2 vaccine1 and 1 vaccine2.
Nurse 3 (pid=861725) has brought vaccine 2: the clinic has 2 vaccine1 and 2 vaccine2.
Vaccinator 2 (pid=861730) is inviting citizen pid=861727 to the clinic.
Citizen 2 (pid=861727) is vaccinated for the 3rd time: the clinic has 1 vaccine1 and 1 vaccine2. The citizen is leaving. Remaining citizens to vaccinate: 1
Vaccinator 1 (pid=861729) is inviting citizen pid=861728 to the clinic.
Citizen 3 (pid=861728) is vaccinated for the 3rd time: the clinic has 0 vaccine1 and 0 vaccine2. The citizen is leaving. Remaining citizens to vaccinate: 0
Nurses have carried all vaccines to the buffer, terminating.
All citizens have been vacinated.
Vaccinator 1 (pid=861729) vaccinated 3 doses. Vaccinator 2 (pid=861730) vaccinated 3 doses. Vaccinator 3 (pid=861731) vaccinated 3 doses. The clinic is now close. Stay healthy.
\end{minted}


\newpage

When log functions are out of the critical section:

\begin{minted}[fontsize=\small,frame=single,framesep=3mm]{text}
./program  -n 3 -v 3 -c 3 -b 11 -t 3 -i input
\end{minted}


\begin{minted}[breaklines,fontsize=\footnotesize,frame=single,framesep=3mm]{text}
Welcome to GTU344 clinic. Number of citizens to vaccinate c=3 with t=3 doeses.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 1 vaccine1 and 0 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 1 vaccine1 and 1 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 2 vaccine1 and 1 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 2 vaccine1 and 2 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 3 vaccine1 and 2 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 3 vaccine1 and 3 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 4 vaccine1 and 3 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 4 vaccine1 and 4 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 5 vaccine1 and 4 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 5 vaccine1 and 5 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 5 vaccine1 and 4 vaccine2.
Vaccinator 1 (pid=866631) is inviting citizen pid=866628 to the clinic.
Citizen 1 (pid=866628) is vaccinated for the 1st time: the clinic has 4 vaccine1 and 4 vaccine2. 
Vaccinator 1 (pid=866631) is inviting citizen pid=866630 to the clinic.
Citizen 3 (pid=866630) is vaccinated for the 1st time: the clinic has 3 vaccine1 and 2 vaccine2. 
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 3 vaccine1 and 3 vaccine2.
Vaccinator 2 (pid=866632) is inviting citizen pid=866629 to the clinic.
Citizen 2 (pid=866629) is vaccinated for the 1st time: the clinic has 4 vaccine1 and 3 vaccine2. 
Nurse 3 (pid=866627) has brought vaccine 1: the clinic has 4 vaccine1 and 3 vaccine2.
Vaccinator 1 (pid=866631) is inviting citizen pid=866628 to the clinic.
Nurse 1 (pid=866625) has brought vaccine 2: the clinic has 3 vaccine1 and 3 vaccine2.
Citizen 1 (pid=866628) is vaccinated for the 2nd time: the clinic has 3 vaccine1 and 2 vaccine2. 
Vaccinator 2 (pid=866632) is inviting citizen pid=866629 to the clinic.
Vaccinator 1 (pid=866631) is inviting citizen pid=866630 to the clinic.
Citizen 3 (pid=866630) is vaccinated for the 2nd time: the clinic has 1 vaccine1 and 1 vaccine2. 
Citizen 2 (pid=866629) is vaccinated for the 2nd time: the clinic has 2 vaccine1 and 2 vaccine2. 
Nurse 3 (pid=866627) has brought vaccine 1: the clinic has 1 vaccine1 and 0 vaccine2.
Nurse 3 (pid=866627) has brought vaccine 2: the clinic has 1 vaccine1 and 1 vaccine2.
Nurse 1 (pid=866625) has brought vaccine 1: the clinic has 2 vaccine1 and 1 vaccine2.
Vaccinator 3 (pid=866633) is inviting citizen pid=866628 to the clinic.
Vaccinator 2 (pid=866632) is inviting citizen pid=866629 to the clinic.
Nurse 3 (pid=866627) has brought vaccine 2: the clinic has 1 vaccine1 and 1 vaccine2.
Citizen 1 (pid=866628) is vaccinated for the 3rd time: the clinic has 0 vaccine1 and 0 vaccine2. The citizen is leaving. Remaining citizens to vaccinate: 2
Citizen 2 (pid=866629) is vaccinated for the 3rd time: the clinic has 1 vaccine1 and 0 vaccine2. The citizen is leaving. Remaining citizens to vaccinate: 1
Vaccinator 2 (pid=866632) is inviting citizen pid=866630 to the clinic.
Citizen 3 (pid=866630) is vaccinated for the 3rd time: the clinic has 0 vaccine1 and 0 vaccine2. The citizen is leaving. Remaining citizens to vaccinate: 0
All citizens have been vacinated.
Nurses have carried all vaccines to the buffer, terminating.
Vaccinator 1 (pid=866631) vaccinated 4 doses. Vaccinator 2 (pid=866632) vaccinated 4 doses. Vaccinator 3 (pid=866633) vaccinated 1 doses. The clinic is now close. Stay healthy.
\end{minted}
\end{document}